---
title: "Project Two"
output:
  pdf_document: default
  html_document: default
---

Due Oct. 13 at 11:59 PM. 

For this first part of the exam, you can either use `surveys_complete.csv` or your own data. If you are using your own data, you must have data in which you think you have a numerical predictor variable and a numerical response variable. If you are using `surveys_complete`, you can use weight and hindfoot_length for this.

Save this file in your `projects` directory. You can either save it in a project two subdirectory, or just put it in projects. Either way is fine.


1) Load in your data. Which variable will you be using as a predictor, and which as a response? (5 pts)

```{r}
library(tidyverse)
surveys <- read_csv("surveys.csv") 
surveys_complete <- na.omit(surveys)

```

```
# The predictor is weight and the response is hindfoot_length
```

2) Plot the two against each other with a scatter plot. Do the data appear to be related linearly? (5 pts)


```{r}
ggplot(surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point()
```

```
# The data is not linear 
```


3) Fit the linear model. View the summary. (5 pts)


```{r}
model_fit <- lm(weight ~ hindfoot_length, data = surveys_complete)
summary(model_fit)
```

4) Does the summary make sense when you compare it to the plot you made? Does our model have good predictive power? Evaluate the residual standard error, intercept, and R-Squared in particular. Would you say your predictor predicts the response?  (10 pts)


```
# The plot and the summary make sense which is why the graph wasn't linear. The R squared being 0.4673 means the predictive power of the weight determining hindfoot length is slightly below average so I wouldn't say it doesn't predict the response very well. The residual standard error was 26 and the closer it is to zero the better fit the model is so that's also indicitive of the model being poor. Also the intercept tells us what we expect the weight to be -32 when the hindfoot length is 2 which also doesn't make much sense as the weight wouldn't be negative. This all goes to say that the predictor doesn't really predict the response well.   
```


5) Plot the model on the graph. Increase the size of the text so it is comfortably readable at 5 feet. Make sure axis labels are readable and not overlapping with one another. (5 pts)

```{r}
ggplot(surveys_complete, aes(x = weight, y = hindfoot_length)) + geom_point() + geom_smooth(method = "lm", size = 0.5) + theme(text = element_text(size = 20))
```


6) Check the normality of the residuals. Do they look OK? Are we violating assumptions? (5 pts)

```{r}

augmented_fit <- broom::augment(model_fit)
qqnorm(augmented_fit$.resid)

# The assumptions were violated because they were not distributed which it was assumed that they were going to be distributed, so the residuals don't look okay. 

```

Why is normality of residuals important? 

```{r}

# The normality of residuals is important to assure the prediction(s) being made are accurate. If they weren't normal then the prediction of the intervals would be inaccurate. 
```

7) If you are using `surveys_complete`: Is there interspecific variation in the linear model? Hint: look at our prior work with facet plots. (15 pts) 

If you are *not* using  `surveys_complete`: Do you think there are groupings within your data that may have a separate linear model? Try grouping the data by that variable and redoing the lm. If this would not make sense for your data, you may also attempt to do multiple predictor variables. (15 pts)

```{r}
ggplot(data = surveys_complete, mapping = aes(x = weight, y = hindfoot_length)) + geom_smooth(method = "lm") + geom_line() + facet_wrap(facets = vars(species_id))

# From looking at the facet plots of the surveys complete data, there does appear to be variation across the species in terms of the weight prediction the hindfoot length. 

```

## Part Two

In this portion, you are free to use your own data if you have a categorical predictor variable and a response variable. Otherwise, you may use the columns sex and weight in `surveys_complete`

1) First, plot the data grouped by sex (5 pts)

```{r}

surveys_complete %>% 
  group_by(sex) %>% 
  ggplot(aes(x = weight, y = hindfoot_length, color = sex)) + geom_point() + scale_color_manual(values = c("pink", "blue"))
```

2) Try an ANOVA of this data (5 pt)

```{r}
model_fit <- lm(hindfoot_length ~ weight, data = surveys_complete)
anova(model_fit)
```

3) How about a linear model? What information does this give you that an ANOVA can't? (5 pts)


```{r}
model_fit <- lm(hindfoot_length ~ weight, data = surveys_complete)
summary(model_fit)
```

```
# The anova doesn't have the r sqaured, residual standard error or the intercept that the linear model does so it can't tell us about its predicitve power or how much deviation there is. 
```

3) Plot the lm with the data, with points colored by sex. (10 pts)


```{r}
surveys_complete %>% 
  group_by(sex) %>% 
  ggplot(aes(x = weight, y = hindfoot_length, color = sex)) + geom_point() + geom_smooth(method = "lm") + scale_color_manual(values = c("pink", "blue"))
```

4) Choose any model we've looked at so far, but use two predictor variables. Perform an LM, plot the results, and state if this changes your interpretation of the relationship between variables (10 pts)

```{r}
model_fit <- lm(hindfoot_length ~ weight + sex, data = surveys_complete)
summary(model_fit)
```

```{r}
surveys_complete %>% 
  group_by(sex) %>% 
  ggplot(aes(x = weight, y = hindfoot_length, color = sex)) + geom_point() + geom_smooth(method = "lm") + scale_color_manual(values = c("pink", "blue"))

```

```
# It basically was saying that weight and sex predict hindfoot length but as it was mentioned before the weight is not a very good predictor of hindfoot length and sex isn't great either. Sex being a predictor as well changes the relationship slightly as its saying that weight and sex are predicting hindfoot_length but it is very similar to the original variables. The only good thing is that the plot is easier to read because of the colors. 
```

## Part Three


1) Add and commit this document (5 pts)

```
# Go to Git in the tool bar area then hit commit and then find projects and unselect everything except project_two and then you can write your commit and hit the commit button
```

2) Push your changes to github (10 pts)

```
# Once you write the commit and exit you can go to the git by connections and tutorial and hit the green push button and it sends it to github
```



# MS students

My expectation is that you'll do this with your own data. If any part of this doesn't make sense with your data, please get in touch ASAP so we can work it out.
